---
title: "repro_research_markdown"
author: "Abigail Kreuser"
date: "1/25/2022"
output: html_document
---

## Things coming up

clone hw repo- start HW
project proposals - want to have that to be ready

#### Notes:

Doesn't use equal signs because cs bad practice 


example: 
``` {r}
y = 2

#versus

x <- 5


```


R is a language with alot of ()

```{r, echo=TRUE, eval=FALSE}
#was saying something about rendering 
rmarkdown::render('yourFile.Rmd')

#lexicalscoping so that you are selecting a certain packages function

```


See comments above about lexical scoping, I have no idea how to spell that

if you load a package R uses the most recently loaded packages function word 
not the base one. 

data sets that come preloaded 


# I MISSED THE CODE WHERE YOU PRINT OUT ATTACHCED PACKAGES 

when thinking about reproducibility, try to minimize the number of attached packages you are working on because future updates and things could. 

tips: teach by typing, exercises are very helpful cementing the fxns and things learned 

```{r}
num <- c(1,2,3,4)
char <- LETTERS[1:4]
fact <- as.factor(c('a','a','b','d'))
logc <- c(TRUE, FALSE, TRUE, TRUE)

# concatenate - creates the vector 
#numeric, character, then there is factor - asigns a numeric value to the variable

fact
#gives levels 

#logical also is called boolean
# treats as numeric 0/1s, good for indexing things -- what is indexing
logc + 0


seq(1,4, by = 1)
#seq is a very useful ... just need ot learn how to use it


is.numeric(num)
#bool response 

str(num)
#str is structure 















```

vectors are useful - but doesnt allowed mixed types?

```{r}
data(mtcars)

head(mtcars)
#it has no name for the type of car 

colnames(mtcars)
rownames(mtcars)

typeof(mtcars)
str(mtcars)
```

So R loaded it as a list, and structure is data.frame

could be a matrix because it is all the same type - numeric 

this matrix is 2D could be more 

wee havent discussed lists yet. we will when we talk about APIs for sure. 

data.frames are cool because they are basically lists, and thats nice for indexing 

```{r}

mtcarsM <- as.matrix(mtcars)
typeof(mtcarsM)
str(mtcarsM)


# the structure fxn gives more insight when mtcars is a data frame
#rownames are sticky when coding more... tbd

```

Now for a short break...
100 rows, 5 columnns 

``` {r}

try3 <- matrix(1:100, nrow = 100, ncol = 5)
try3

oops <- matrix(rows, nrow = 100, ncol = 5)
oops


rows = c(1:100)

col = as.list(1:5)
print(col)

test1 <- matrix(col, rows)
head(test1)
rows
?col

test2 <- col(rows)

```



```{r}
testList <- list(a=1:100, b=c('A', 'B'), d=oops)
str(testList)

#indexing
num[1]

testList[1]

testList[1]+5 # no work
testList[[1]]+5
typeof(testList[1])
newlist <-testList[1:2]
newlist

mtcars$cyl
mtcars[['cyl']]


```

# Review notes 

class vs typeof vs str give all different insights

```{r}

#remove things from matrix using [-1,]
# ways to look at the end of data not in hard code 


nrow(oops)

tail(oops, 1)

oops[nrow(oops),]

test2 = rbind(oops, oops) #row bind
cbind(oops, oops) #binding by columns

testDF <- as.data.frame(test2)

oops[30,4]
str(testDF)
testDF$V4[30]




```
if{all,(is.na(test[,1]))} 
or 
if{any(is.na(test[,1]))}

im lost... !

Conditionals 

2022-02-01
Picking up here 


```{r}


unique(test[,2])

brody <- 'cat'
brody == 'cat'
brody != 'cat'
brody == 'dog'


animals <- c('dog', 'cat', 'rhino', NA )

brody == animals


sum(na.omit(c(1:10,NA)))

# can use na.rm= TRUE
# or na.omit



###########################  
#Uh move this
#index all of the rows in the matrix whose first column value 2 
test = matrix(runif(500), ncol=5)
# remove rows whose first colum is less than .5


point5 <- test[,1]> 0.5 
a <- test[test[,1]> 0.5, ] 
b <- test[point5,]
dim(b)
sum(point5)
a == b
all(a == b)

which(point5)

# A VERY POWERFULL TOOL %in%

```

NA values are different then NULL values. 
there can only be 1 NULL and it can be weird and throw or not throw errors


``` {r}

is.na(animals)

is.null(animals)

c(NULL, NULL, NULL) + 1
Inf-Inf #NaN is not a number 

all(brody == animals)
any(brody == animals)
brody %in% animals

which(brody == animals)
# this gives you the index number and then can use that if needed 


```

if (condition){
(action)
}else{
(other action)
}
  
```{r}
if (any(brody == animals)){
  print('brody is an animal')
}


if (any(brody == animals)){
  print('brody is an animal')
}else{
  print("brody is not an animal")
}





```
  If else gives a binary example, but could need a string of ifs and nested ifs and Future examples 
  
```{r}
  
#LOOK AT THE YOUTUBES 

mean(unlist(arizona2[,12:ncol(arizona2)]))
```


LOOPS

loops are helpful not silly 
only silly if you nest like 3 loops 

for every i where i = a vector or numeric index of my vegtables 

```{r}

a <- 1:10
b <- 10:1

# vectorized version 
d <- a + b 
#answer is 10 11's
d <- c()
for(i in 1:length(a)){
  d[i] <- a[i] + b[i]
}

# break this down by running 
i = 1
i = 2
i = 5 

# this will assign 11s to the 1, 2, 5th position in d, and leave the others as NA


testList2 <- list(a=runif(100), b=runif(5), d=runif(1000))

out <- c()


#length a because were going as long as length a 



```